---
output:
  html_document:
    theme:
      version: "4"
      bg: "#FFFFFF"
      fg: "#072B67"
      primary: "#C2326B"
    df_print: paged
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(tidyverse)
```

# Harmonized Historical Census Data

<!-- badges: start -->
<!-- badges: end -->

## Overview

The U.S. Census is a rich source for population-level demographic, socioeconomic, and housing data. Although the Census Bureau has been collecting data since the 1700s, the data collected has changed over time, with topics being removed, added, and redefined decade-to-decade, making long-term longitudinal analyses difficult. Here, we sought to compile and harmonize variables from the 1970 census through the 2010 census at the census tract level in the following categories:

- population
- income
- poverty
- assisted income
- unemployment
- educational attainment
- racial composition
- housing occupancy
- housing value

## Getting the Data

### Downloading the CSV file

The data is contained in a CSV file called [harmonized_census_data.csv](https://github.com/geomarker-io/harmonized_historical_census_data/blob/main/harmonized_census_data.csv) which is a table of all census tracts listed by their FIPS ID and corresponding  .

### Import Directly Into `R`

Use the following code to download the deprivation index data.frame directly into R:

```
census_data <- 'https://github.com/geomarker-io/harmonized_historical_census_data/blob/main/harmonized_census_data.rds' %>% 
    url() %>% 
    gzcon() %>% 
    readRDS() %>% 
    as_tibble()
```

## Usage

### Matching longitudinal address data to census tracts

First, geocode longitudinal address data using the DeGAUSS [geocoder](https://degauss.org/geocoder/).

Then, use the DeGAUSS [Spatiotemporal Census Tract container](https://degauss.org/st_census_tract/) to assign census tract identifiers for the appropriate decade. This container requires a date range (`start_date` and `end_date` columns). Rows with date ranges that span mulitple decades will be split to one row per decade. 

For example, 

```{r, echo = FALSE, message = FALSE}
read_csv('example_data/my_address_file_geocoded_split_ex.csv')
```

would become 

```{r, echo = FALSE, message = FALSE}
address_data <- read_csv('example_data/my_address_file_geocoded_split_ex_st_census_tract_v0.1.3.csv', 
                         col_types = cols(census_tract_vintage = col_character(), 
                                          census_tract_id = col_character()))
address_data
```

where a 2010 tract identifier is assigned to the first row, and the second row is **split** into one row with a **2000** tract identifier and one row with a **2010** tract identifier.

### Joining to harmonized_historical_census_data

Once `census_tract_vintage` and `census_tract_id` is assinged for each row, the `harmonized_census_data` can be joined to your longitudinal address data using both the `census_tract_vintage` and `census_tract_id` columns as keys. 

For example, in `R` this could look like

```{r, echo = FALSE, message = FALSE}
harmonized_census_data <- readRDS('harmonized_census_data.rds')
```

```{r}
dplyr::left_join(address_data, harmonized_census_data, by = c("census_tract_vintage", "census_tract_id"))
```

## Details

The table below includes specific variable definitions for each census decade.

```{r, echo = FALSE, message = FALSE}
dd <- readxl::read_excel('harmonized_census_data_dictionary.xlsx', skip = 1) %>% 
  select(`Harmonized Variable Name`, 
         `Harmonized Description` = Description...2,
         `1970 NHGIS dataset` = `NHGIS dataset...3`,
         `1970 Derivation` = `Census Variables...4`,
         `1970 Description` = Description...5,
         `1980 NHGIS dataset` = `NHGIS dataset...6`,
         `1980 Derivation` = `Census Variables...7`,
         `1980 Description` = Description...8, 
         `1990 NHGIS dataset` = `NHGIS dataset...9`,
         `1990 Derivation` = `Census Variables...10`,
         `1990 Description` = Description...11,
         `2000 NHGIS dataset` = `NHGIS dataset...12`,
         `2000 Derivation` = `Census Variables...13`,
         `2000 Description` = Description...14,
         `2010 NHGIS dataset` = `NHGIS dataset...15`,
         `2010 Derivation` = `Census Variables...16`,
         `2010 Description` = Description...17)

DT::datatable(
  dd, 
  width = 1100,
  extensions = c('Scroller', 'Buttons'),
  options = list(
    deferRender = TRUE,
    scrollY = 1000,
    scrollX = 200,
    scroller = TRUE,
    dom = 'Bfrtip',
    buttons = c(list(list(extend = 'colvis', columns = c(3:17))), 
                list(list(extend = 'collection',
                  buttons = c('csv', 'excel'),
                  text = 'Download as')))
))
```

## Reproducibility

All census data was obtained from [nhgis.org](https://www.nhgis.org/). Each `harmonized_historical_census_data` variable was derived using the formulas listed under the `derivation` columns, and census variables included in each formula column can be found in the NHGIS dataset listed in the corresponding `NHGIS dataset` column. 


