---
output:
  html_document:
    theme:
      version: "4"
      bg: "#FFFFFF"
      fg: "#072B67"
      primary: "#C2326B"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(tidyverse)
```

# Harmonized Historical Census Data

<!-- badges: start -->
<!-- badges: end -->

## Overview

The U.S. Census is a rich source for population-level demographic, socioeconomic, and housing data. Although the Census Bureau has been collecting data since the 1700s, the data collected has changed over time, with topics being removed, added, and redefined decade-to-decade, making long-term longitudinal analyses difficult. Here, we sought to compile and harmonize variables from the 1970 census through the 2010 census at the census tract level in the following categories:

- population
- income
- poverty
- assisted income
- unemployment
- educational attainment
- racial composition
- housing occupancy
- housing value

## Getting the Data

### Downloading the CSV file

The data is contained in a CSV file called [harmonized_census_data.csv](https://github.com/geomarker-io/harmonized_historical_census_data/blob/main/harmonized_census_data.csv) which is a table of all census tracts listed by their FIPS ID and corresponding  .

### Import Directly Into `R`

Use the following code to download the deprivation index data.frame directly into R:

```
census_data <- 'https://github.com/geomarker-io/harmonized_historical_census_data/blob/main/harmonized_census_data.rds' %>% 
    url() %>% 
    gzcon() %>% 
    readRDS() %>% 
    as_tibble()
```

## Usage

Suppose you have a ...

### Matching longitudinal address data to census tracts

- geocoder
- st_census_tract

### Joining 

- by census_tract_vintage and census_tract_id


## Details

The table below includes specific variable definitions for each census decade.

```{r, echo = FALSE, message = FALSE}
dd <- readxl::read_excel('harmonized_census_data_dictionary.xlsx', skip = 1) %>% 
  select(`Harmonized Variable Name`, 
         `Harmonized Description` = Description...2,
         `1970 NHGIS dataset` = `NHGIS dataset...3`,
         `1970 Derivation` = `Census Variables...4`,
         `1970 Description` = Description...5,
         `1980 NHGIS dataset` = `NHGIS dataset...6`,
         `1980 Derivation` = `Census Variables...7`,
         `1980 Description` = Description...8, 
         `1990 NHGIS dataset` = `NHGIS dataset...9`,
         `1990 Derivation` = `Census Variables...10`,
         `1990 Description` = Description...11,
         `2000 NHGIS dataset` = `NHGIS dataset...12`,
         `2000 Derivation` = `Census Variables...13`,
         `2000 Description` = Description...14,
         `2010 NHGIS dataset` = `NHGIS dataset...15`,
         `2010 Derivation` = `Census Variables...16`,
         `2010 Description` = Description...17)

DT::datatable(
  dd, 
  extensions = c('Scroller', 'Buttons'),
  options = list(
    deferRender = TRUE,
    scrollY = 200,
    scrollX = 200,
    scroller = TRUE,
    dom = 'Bfrtip',
    buttons = c(list(list(extend = 'colvis', columns = c(3:17))), 
                list(list(extend = 'collection',
                  buttons = c('csv', 'excel'),
                  text = 'Download as')))
))
```

[harmonized_census_data_dictionary.xlsx](https://github.com/geomarker-io/harmonized_historical_census_data/blob/main/harmonized_census_data_dictionary.xlsx).

## Reproducibility

- steps ?


